---
title: Lab Eleven -- Point Estimation and Resampling # Title of the document

# Bibliography Details
bibliography: Lab11References.bib  # Name of bib file in the same folder 
bibliographystyle: apa            # Bibliography style

execute:               # Default settings for code blocks
  warning: false       # Do not print warnings
  message: false       # Do not print messages
  size: "scriptsize"   # Set code size to script size

# Below will do html by default. To render the entire project
# Use the following code:
# quarto::quarto_render("Lab10WriteUp.qmd")
format:
  pdf:        # Specify .pdf formatting options 
    documentclass: article     # Document Class
    number-sections: true      # Number Sections
    reference-section-title: "References" # Title for reference section
    code-line-numbers: true    # Line numbers printed in code
    cap-location: top          # Captions at the top of tables 
    geometry:
      - top=0.75in
      - left=0.75in
      - bottom=0.75in
      - right=0.75in
  html:       # Specify HTML formatting options 
    number-sections: true      # Number Sections
    code-fold: true            # Button to show code
    code-line-numbers: true    # Line numbers printed in code
    cap-location: top          # Captions at the top of tables and figures
format-links: [pdf]
---
```{r}
#| echo: false

default_chunk_hook  <- knitr::knit_hooks$get("chunk")

latex_font_size <- c("Huge", "huge", "LARGE", "Large", 
                     "large", "normalsize", "small", 
                     "footnotesize", "scriptsize", "tiny")

knitr::knit_hooks$set(chunk = function(x, options) {
  x <- default_chunk_hook(x, options)
  if(options$size %in% latex_font_size) {
    paste0("\n \\", options$size, "\n\n", 
      x, 
      "\n\n \\normalsize"
    )
  } else {
    x
  }
})
```

```{r}
#| label: load-packages
#| include: false
library(tidyverse)
library(patchwork)
```

**Tips**

* Complete the tasks below. Make sure to start your solutions in on a new line that starts with "**Solution**:".
* Make sure to use the Quarto Cheatsheet. This will make completing and writing up the lab *much* easier.
* Make sure to use "enter" to make your code readable, so it doesn't run off the page. I switched the Quarto document to automatically render to .pdf, so you can see what you're submitting by default. You can switch back for the highlighting by moving the html block above the pdf block in the YAML header.
* Don't forget that you can copy-and-paste code when you're doing something similar as we did in class!
* Make sure to plot all computed probabilities. This is good `ggplot` practice AND will help make sure you don't make a mistake when computing the probability.


# Introduction
When performing point estimation, there isn't just one *right* answer. For example, the Method of Moments (MOM) and Maximum Likelihood Estimates (MLEs) aren't always the same.

One question that comes up, is why we calculate the sample variance as
$$S_{n-1}^2 = \frac{1}{n-1} \sum_{i=1}^n (X_i -\bar{X})^2,$$
when it is supposed to be the ``average squared distance from the mean, i.e., 
$$S_n^2 = \frac{1}{n} \sum_{i=1}^n (X_i -\bar{X})^2.$$


# Comparing Estimators

## Initial Simulation
In this lab, we will explore why we use $S_{n-1}^2$ and not $S_n^2$. To do this, we will generate data from four known distributions:

* Poisson$(\lambda=2)$
* Binomial$\left(n=50, p=\frac{5-\sqrt{21}}{10}\right)$
* Normal$(\mu=0, \sigma=\sqrt{2})$
* Exponential$\left(\lambda = \frac{1}{\sqrt{2}}\right).$

**Remark**: Note that I have selected the parameters so that the true variance for all four models is 2.

For each distribution, write a loop that completes the following 1000 times. Put `set.seed(7272)` at the top of your code chunk.

* Generate a sample of size $n$ from the distribution.
* Calculate and save $S_{n-1}^2$ and $S_n^2.$

Start with $n=20$. Is there evidence that $S_{n-1}^2$ is a better estimator than $S_n^2$?


Create a plot or combination of plots (using patchwork) to show the results across all four distributions. Make sure to create a corresponding table that numerically summarizes the results of the simulation. You should include
\begin{align*}
\textrm{Bias} &= \textrm{Average of Estimates} - \textrm{True Value}\\
\textrm{Precision} &= \frac{1}{\textrm{Variance of Estimates}}\\
\textrm{Mean Square Error} &= \textrm{Bias}^2 + \textrm{Variance of Estimates}.
\end{align*}

**Solution**
```{r}
set.seed(7272)

# Poisson
P.S.estimate = rep(NA, 1000)
P.Sn.estimate = rep(NA, 1000)

for(i in 1:1000){
  X <- rpois(20, lambda = 2)
  S <- var(X)
  Sn <- 19/20 * var(X)
  P.S.estimate[i] <- S
  P.Sn.estimate[i] <- Sn
}

# Binomial
B.S.estimate = rep(NA, 1000)
B.Sn.estimate = rep(NA, 1000)
P = (5 - (21^0.5))/10
for(i in 1:1000){
  X <- rbinom(20, 50, P)
  S <- var(X)
  Sn <- 19/20 * var(X)
  B.S.estimate[i] <- S
  B.Sn.estimate[i] <- Sn
}

# Normal
N.S.estimate = rep(NA, 1000)
N.Sn.estimate = rep(NA, 1000)

for(i in 1:1000){
  X <- rnorm(20, 0, (2^0.5))
  S <- var(X)
  Sn <- 19/20 * var(X)
  N.S.estimate[i] <- S
  N.Sn.estimate[i] <- Sn
}

# Exponential
E.S.estimate = rep(NA, 1000)
E.Sn.estimate = rep(NA, 1000)

for(i in 1:1000){
  X <- rexp(20, 1/(2^0.5))
  S <- var(X)
  Sn <- 19/20 * var(X)
  E.S.estimate[i] <- S
  E.Sn.estimate[i] <- Sn
}

results <- tibble(PoissonS = P.S.estimate,
                 PoissonSn = P.Sn.estimate,
                 BinomialS = B.S.estimate,
                 BinomialSn = B.Sn.estimate,
                 NormalS = N.S.estimate,
                 NormalSn = N.Sn.estimate,
                 ExponentialS = E.S.estimate,
                 ExponentialSn = E.Sn.estimate)

results.long <- results |>
  pivot_longer(cols = everything(),names_to = "Estimator", values_to = "Estimate")

summary <- results.long |>
  group_by(Estimator) |>
  summarize(mean = mean(Estimate),
            variance = var(Estimate)) |>
  mutate(bias = mean-2,
        precision = 1/variance,
        mse = variance + bias^2)

Poisson <- results.long |>
  filter(Estimator == "PoissonS" | Estimator == "PoissonSn")

Poisson.plot <- ggplot(Poisson) +
  stat_density(aes(x=Estimate, color=Estimator), geom="line", position = "jitter") +
  geom_vline(xintercept=2, color="red", linetype="dotted")+
  geom_hline(yintercept=0)+
  theme_bw()+
  labs(x=bquote(hat(S^2)), y="Density")

Binomial <- results.long |>
  filter(Estimator == "BinomialS" | Estimator == "BinomialSn")

Binomial.plot <- ggplot(Binomial) +
  stat_density(aes(x=Estimate, color=Estimator), geom="line", position = "jitter") +
  geom_vline(xintercept=2, color="red", linetype="dotted")+
  geom_hline(yintercept=0)+
  theme_bw()+
  labs(x=bquote(hat(S^2)), y="Density")

Normal <- results.long |>
  filter(Estimator == "NormalS" | Estimator == "NormalSn")

Normal.plot <- ggplot(Normal) +
  stat_density(aes(x=Estimate, color=Estimator), geom="line", position = "jitter") +
  geom_vline(xintercept=2, color="red", linetype="dotted")+
  geom_hline(yintercept=0)+
  theme_bw()+
  labs(x=bquote(hat(S^2)), y="Density")

Exponential <- results.long |>
  filter(Estimator == "ExponentialS" | Estimator == "ExponentialSn")

Exponential.plot <- ggplot(Exponential) +
  stat_density(aes(x=Estimate, color=Estimator), geom="line", position = "jitter") +
  geom_vline(xintercept=2, color="red", linetype="dotted")+
  geom_hline(yintercept=0)+
  theme_bw()+
  labs(x=bquote(hat(S^2)), y="Density")

Poisson.plot 
Binomial.plot
Normal.plot 
Exponential.plot
```

```{r}
#| label: summary1
#| tbl-cap: "Summary of the result of the simulation"
#| tbl-align: "center"
#| echo: false
summary |>
  knitr::kable(digits = 2)
```

## Simulation over various sample sizes
Consider the effect of the sample size have on these results. Provide a graphical and numerical summary of what happens as $n$ increases -- say $n=20$, $n=50$, $n=100$, and $n=500$. Does the result in your initial simulation hold across samples? Or, do things change as the sample size changes? Ensure to set your randomization seed.
 
## Simulation
You have written code to draw a random sample of various sizes from each distribution, and to calculate and store the calculated estimates. Rework your code to produce a density plot of the $s_{n-1}^2$ and $s_{n}^2$. Do you have a guess about their distribution? Ensure to set your randomization seed. 

## Bootstrap
Now, take *one* sample from each distribution. Instead of drawing a random sample at each iteration, sample *with replacement* from the one sample. Produce a density plot of the $s_{n-1}^2$ and $s_{n}^2$. Do we get roughly the same information as we do in the full simulation? Try this over several seeds -- what do you notice? Ensure to set your randomization seed.


# Executive Summary
Summarize the work you've done here to provide a brief (200-300 word) summary of the simulation and the results using your work in Lab 11. Your summary should be professional, clear, and all claims should be corroborated with statistical support.